---
title: "Introduction to mypaintr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# mypaintr

mypaintr is a package for creating artistic sketch-like plots in R. It has
three components:

* An R interface to the [libmypaint](https://github.com/mypaint/libmypaint) 
  library, which lets you create and import Mypaint brushes. There's a
  `mypaint_device()` graphics device.
* R functions to draw lines and shapes with a "rough", hand-drawn look.
* ggplot geoms and theme elements for ggplot graphs using Mypaint brushes 
  and hand-drawn paths.
  
Here are some demos.

```{r setup, include = FALSE}
library(mypaintr)
knitr::knit_hooks$set(mypaint = knitr_chunk_hook())

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.keep = "none",
  fig.ext = "png",
  mypaint = TRUE,
  fig.width = 5
)
```

Brushes are from the mypaint-brushes package which you can install via
`brew` or your package manager.

```{r}
set_hand(hand(seed = 1))
set_brush("deevad/2B_pencil")
barplot(VADeaths, beside = TRUE, col = palette.colors(5), cex.names = 0.8)
```


```{r}
set_hand(NULL)
set_brush("experimental/bubble")
barplot(VADeaths, beside = TRUE, col = palette.colors(5), cex.names = 0.8)
```

