---
title: "Introduction to mypaintr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


mypaintr is a package for creating artistic sketch-like plots in R. It has
three components:

* An R interface to the [libmypaint](https://github.com/mypaint/libmypaint) 
  library, which lets you create and import Mypaint brushes. There's a
  `mypaint_device()` graphics device.
* R functions to draw lines and shapes with a "rough", hand-drawn look.
* ggplot geoms and theme elements for ggplot graphs using Mypaint brushes 
  and hand-drawn paths.
  
Here are some demos.

```{r setup, include = FALSE}
library(mypaintr)
knitr::knit_hooks$set(mypaint = knitr_chunk_hook())

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.ext = "png",
  fig.width = 5
)
```

Brushes are from the mypaint-brushes package which you can install via
`brew` or your package manager.

```{r, mypaint=TRUE, fig.keep="none"}
set_hand(hand(seed = 1))
set_brush("tanda/marker-01")
barplot(VADeaths, beside = TRUE, col = NA, cex.names = 0.8)
```

Set `hand` or `brush` to `NULL` to revert to the default:

```{r, mypaint=TRUE, fig.keep="none"}
set_hand(NULL)
set_brush("ramon/P._shade")
barplot(VADeaths, axes = FALSE, 
        beside = TRUE, col = palette.colors(5), border = NA,
        cex.names = 0.8)
set_brush(NULL)
axis(side = 2, at = seq(0, 60, 20))
```

# Rough lines and polygons

These functions can be used with any graphics device, as in the next chunks:

```{r, dev="png"}
knitr::opts_chunk$get("dev")

plot(1:10, 1:10, type = "n")
draw_rough_polygons(c(2, 4, 6), c(4, 2, 6), col = "red")
draw_rough_rect(8, 4, 5, 8, density = 15, col = "blue3")

draw_rough_arrows(1, 9, 8, 9, col = "grey40")
```

Control lines and fills with the `hand()` function:


```{r, dev="png"}

plot(1:10, 1:10, type = "n")

my_hand <- hand(wobble = 0.01, multi_stroke = 2)
draw_rough_polygons(c(2, 4, 6), c(4, 2, 6), col = "red", hand = my_hand)
draw_rough_rect(8, 4, 5, 8, density = 15, col = "blue3", hand = my_hand)

draw_rough_arrows(1, 9, 8, 9, col = "grey40", hand = my_hand)
```



```{r, dev="png", fig.width=5, fig.height=5}

plot(5:55, 5:55, type = "n", xlab = "bow", ylab = "wobble", axes = FALSE,
     mar = rep(0.1, 4))

for (row in 1:5) for (col in 1:5) {
  my_hand <- hand(wobble = 0.02 * row, bow = 0.02 * col)
  draw_rough_rect(
    10 * col - 3, 10 * row - 3,
    10 * col + 3, 10 * row + 3,
    hand = my_hand,
    col = "red"
  )
}
```
